PANCHANGE_data_preprocessing

### 2020-09-02 PANCHANGE_analysis.Rmd

## COPING
Filter data frame to exclude participants that never answered COPING
```{r exclusion criteria dat}
dat <- dat.raw %>%
  filter(
    (data_group_gad == "Baseline and retrospective data" |
      data_group_gad == "Prepandemic and baseline data" |
      data_group_gad == "All prepandemic, baseline and retrospective data") |
      (data_group_phq == "Baseline and retrospective data" |
      data_group_phq == "Prepandemic and baseline data" |
      data_group_phq == "All prepandemic, baseline and retrospective data") |
      (data_group_pcl == "Baseline and retrospective data" |
      data_group_pcl == "Prepandemic and baseline data" |
      data_group_pcl == "All prepandemic, baseline and retrospective data") |
      (data_group_ocir == "Baseline and retrospective data" |
      data_group_ocir == "Prepandemic and baseline data" |
      data_group_ocir == "All prepandemic, baseline and retrospective data")
  )
```

>>>>

## COPING
Filter data frame to exclude participants that never answered COPING
```{r exclusion criteria dat}
dat.2.timepoints <- dat.raw %>%
  filter(
    (data_group_gad == "Baseline and retrospective data" |
      data_group_gad == "Prepandemic and baseline data" |
      data_group_gad == "All prepandemic, baseline and retrospective data") |
      (data_group_phq == "Baseline and retrospective data" |
      data_group_phq == "Prepandemic and baseline data" |
      data_group_phq == "All prepandemic, baseline and retrospective data") |
      (data_group_pcl == "Baseline and retrospective data" |
      data_group_pcl == "Prepandemic and baseline data" |
      data_group_pcl == "All prepandemic, baseline and retrospective data") |
      (data_group_ocir == "Baseline and retrospective data" |
      data_group_ocir == "Prepandemic and baseline data" |
      data_group_ocir == "All prepandemic, baseline and retrospective data")
  )

dim(dat.2.timepoints)
```

```{r}
dat.2.timepoints %>%
  select(age_category) %>%
  filter(is.na(age_category))
```

Exclusion
```{r Exclusion}
dat <- dat.2.timepoints %>%

### 2020-09-02 PANCHANGE_data_preprocessing2020-07-15.Rmd

edgi.mhd.raw.id <- edgi.mhd.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)) %>%
  select(
    Sample, # Sample
    ID, # ID
    mhd.anorexia_nervosa,
    mhd.atypical_anorexia_nervosa,
    mhd.bulimia_nervosa,
    mhd.atypical_bulimia_nervosa,
    mhd.bingeeating_disorder,
    mhd.atypical_bingeeating_disorder,
    mhd.none_of_the_above_eating_disorder = mhd.none_of_the_above
  ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

  edgi.mhd2.raw.id <- edgi.mhd2.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID)) %>%
    select(
      Sample, # Sample
      ID, # ID
    mhd.depression = demographics.depression,
    mhd.pregnancy_depression = demographics.pregnancy_depression,
    mhd.premenstrual_dysphoric_disorder_pmdd = demographics.premenstrual_dysphoric_disorder_pmdd,
    mhd.anxiety_nerves_or_generalised_anxiety_disorder = demographics.anxiety_nerves_or_generalised_anxiety_disorder,
    mhd.social_anxiety_or_social_phobia = demographics.social_anxiety_or_social_phobia,
    mhd.specific_phobia_e.g._phobia_of_flying = demographics.specific_phobia_e.g._fear_of_flying, # fear instead of phobia
    mhd.agoraphobia = demographics.agoraphobia,
    mhd.panic_disorder = demographics.panic_disorder,
    mhd.panic_attacks = demographics.panic_attacks,
    mhd.obsessivecompulsive_disorder_ocd = demographics.obsessivecompulsive_disorder_ocd,
    mhd.obsessive_compulsive_related_disorders = demographics.obsessive_compulsive_related_disorders,
    mhd.body_dysmorphic_disorder_bdd = demographics.body_dysmorphic_disorder,
    mhd.mania_hypomania_bipolar_or_manicdepression = demographics.mania_hypomania_bipolar_or_manicdepression,
    mhd.schizophrenia = demographics.schizophrenia,
    mhd.type_psychosis_psychotic_illness = demographics.psychosis_type_psychotic_illness, # word type psychosis changed order compared to GLAD 
    mhd.personality_disorder_diagnosed = demographics.diagnosed_personality_disorder, # named differently in EDGI
    mhd.posttraumatic_stress_disorder_ptsd = demographics.posttraumatic_stress_disorder_ptsd,
    mhd.autism_aspergers_or_autistic_spectrum_disorder = demographics.autism_spectrum_disorder_asd, # different name in EDGI
    mhd.body_dysmorphic_disorder_bdd = demographics.body_dysmorphic_disorder,
    mhd.personality_disorder = demographics.personality_disorder, # routing variable in EDGI is named differently; renamed here to match GLAD
    mhd.attention_deficit_hyperactivity_disorder = demographics.attention_deficit_hyperactivity_disorder,
    mhd.none_of_the_above = demographics.none_of_the_above.1,
    mhd.none_of_the_above.1 = demographics.none_of_the_above.8
    ) %>%
    add_column(
      mhd.schizoaffective_disorder = NA_character_
    ) %>%
    add_numeric(., exclude = exclude_cols_mhd2) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

      case_when(mhd.suspected_eating_disorder_diagnosed_numeric == "0" ~ "Not Other eating disorder",
                mhd.other_eating_disorder_unc_numeric == "0" ~ "Not Other eating disorder",
                mhd.other_eating_disorder_unc_numeric == "1" ~ "Other eating disorder")

@@

```{r mhd check none of the above}
if(NBR == TRUE) {
mhd.raw.id %>%
  group_by(control) %>%
  count(
    mhd.none_of_the_above.1,
    mhd.none_of_the_above,
    mhd.other_please_tell_us_more,
    mhd.other_eating_disorder
  )
}
```

@@

```{r Investigating RAMP NAs}
dat %>%
  group_by(control) %>%
  count(
    mhd.none_of_the_above.1,
    mhd.none_of_the_above,
    control,
    age_category
  )

@@

dat %>%
  filter(
    (mhd.none_of_the_above.1 == "Not None of these" |
    mhd.none_of_the_above == "Not None of these") &
      control == "Control"
  )

>>>>

edgi.mhd.raw.id <- edgi.mhd.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)) %>%
  select(
    Sample, # Sample
    ID, # ID
    mhd.anorexia_nervosa,
    mhd.atypical_anorexia_nervosa,
    mhd.bulimia_nervosa,
    mhd.atypical_bulimia_nervosa,
    mhd.bingeeating_disorder,
    mhd.atypical_bingeeating_disorder,
    mhd.other_eating_disorder,
    mhd.none_of_the_above_eating_disorder = mhd.none_of_the_above,
    mhd.dont_know_eating_disorder = mhd.dont_know,
    mhd.prefer_not_to_answer_eating_disorder = mhd.prefer_not_to_answer
  ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

  edgi.mhd2.raw.id <- edgi.mhd2.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID)) %>%
    select(
      Sample, # Sample
      ID, # ID
    mhd.depression = demographics.depression,
    mhd.pregnancy_depression = demographics.pregnancy_depression,
    mhd.premenstrual_dysphoric_disorder_pmdd = demographics.premenstrual_dysphoric_disorder_pmdd,
    mhd.anxiety_nerves_or_generalised_anxiety_disorder = demographics.anxiety_nerves_or_generalised_anxiety_disorder,
    mhd.social_anxiety_or_social_phobia = demographics.social_anxiety_or_social_phobia,
    mhd.specific_phobia_e.g._phobia_of_flying = demographics.specific_phobia_e.g._fear_of_flying, # fear instead of phobia
    mhd.agoraphobia = demographics.agoraphobia,
    mhd.panic_disorder = demographics.panic_disorder,
    mhd.panic_attacks = demographics.panic_attacks,
    mhd.obsessivecompulsive_disorder_ocd = demographics.obsessivecompulsive_disorder_ocd,
    mhd.obsessive_compulsive_related_disorders = demographics.obsessive_compulsive_related_disorders,
    mhd.body_dysmorphic_disorder_bdd = demographics.body_dysmorphic_disorder,
    mhd.mania_hypomania_bipolar_or_manicdepression = demographics.mania_hypomania_bipolar_or_manicdepression,
    mhd.schizophrenia = demographics.schizophrenia,
    mhd.type_psychosis_psychotic_illness = demographics.psychosis_type_psychotic_illness, # word type psychosis changed order compared to GLAD 
    mhd.personality_disorder_diagnosed = demographics.diagnosed_personality_disorder, # named differently in EDGI
    mhd.posttraumatic_stress_disorder_ptsd = demographics.posttraumatic_stress_disorder_ptsd,
    mhd.autism_aspergers_or_autistic_spectrum_disorder = demographics.autism_spectrum_disorder_asd, # different name in EDGI
    mhd.body_dysmorphic_disorder_bdd = demographics.body_dysmorphic_disorder,
    mhd.personality_disorder = demographics.personality_disorder, # routing variable in EDGI is named differently; renamed here to match GLAD
    mhd.attention_deficit_hyperactivity_disorder = demographics.attention_deficit_hyperactivity_disorder,
    mhd.none_of_the_above = demographics.none_of_the_above.1,
    mhd.dont_know = demographics.dont_know,
    mhd.prefer_not_to_answer = demographics.prefer_not_to_answer.3,
    mhd.none_of_the_above.1 = demographics.none_of_the_above.8,
    mhd.dont_know.1 = demographics.dont_know.7,
    mhd.prefer_not_to_answer.1 = demographics.prefer_not_to_answer.11,
    mhd.other_please_tell_us_more = demographics.other.1
    ) %>%
    add_column(
      mhd.schizoaffective_disorder = NA_character_
    ) %>%
    add_numeric(., exclude = exclude_cols_mhd2) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

      case_when(mhd.suspected_eating_disorder_diagnosed_numeric == "0" ~ "Not Other eating disorder:",
                mhd.other_eating_disorder_unc_numeric == "0" ~ "Not Other eating disorder:",
                mhd.other_eating_disorder_unc_numeric == "1" ~ "Other eating disorder:")

@@

```{r mhd check none of the above}
if(NBR == TRUE | EDGI == TRUE) {
mhd.raw.id %>%
  group_by(control) %>%
  count(
    mhd.none_of_the_above.1,
    mhd.none_of_the_above,
    mhd.other_please_tell_us_more,
    mhd.other_eating_disorder
  )
}
```

@@

```{r Investigating RAMP NAs}
if(RAMP == TRUE) {
dat %>%
  group_by(control) %>%
  count(
    mhd.none_of_the_above.1,
    mhd.none_of_the_above,
    control,
    age_category
  )

@@

dat %>%
  filter(
    (mhd.none_of_the_above.1 == "Not None of these" |
    mhd.none_of_the_above == "Not None of these") &
      control == "Control"
  )
}
```
